\documentclass[11pt]{letter}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}

\signature{Taesung Park}
\address{Independent Researcher \\ dlfmadldlfmaek@naver.com}

\begin{document}

\begin{letter}{Editorial Board \\ PLOS Computational Biology}

\opening{Dear Editors,}

We are pleased to submit our manuscript entitled ``\textbf{Protein Language Model Scale Determines the Direction of Structure-Aggregation Coupling in Intrinsically Disordered Proteins}'' for consideration as a Research Article in \textit{PLOS Computational Biology}.

\textbf{Summary.} We present a conditional flow matching framework that generates mutation-specific conformational ensembles for intrinsically disordered proteins (IDPs) and predicts their aggregation propensity. Using deep mutational scanning data for two amyloid peptides---amyloid-$\beta$ 42 ($n$=751) and IAPP ($n$=661)---we demonstrate that the change in radius of gyration ($\Delta R_g$) from flow-generated ensembles significantly predicts experimental nucleation scores, outperforming the recently published CANYA predictor (Science Advances, 2025).

\textbf{Key finding.} We discover an unexpected qualitative phase transition in protein language model (PLM) representations: the direction of the structure--aggregation correlation reverses between ESM-2 150M and 650M, even when the downstream model architecture is held constant. This sign change---not merely a change in magnitude---has not been previously reported in the PLM scaling literature and reveals that larger language models encode fundamentally different structural information about IDP conformational dynamics.

\textbf{Significance.} This work is relevant to the readership of \textit{PLOS Computational Biology} because it:

\begin{itemize}
    \item Establishes flow matching as a lightweight, MD-free alternative for mutation-specific IDP ensemble generation
    \item Provides the first multi-amyloid validation (A$\beta$42 + IAPP) of a generative structure model for aggregation prediction, including out-of-distribution generalization
    \item Reveals scale-dependent behavior in protein language models with implications beyond IDP research for the broader PLM community
    \item Reports all results with bootstrap confidence intervals and multiple comparison correction, with full code and data publicly available
\end{itemize}

All code is available at \url{https://github.com/layeredlabs06/brain-idp-flow}, and a Google Colab notebook reproduces all results.

This manuscript has not been published elsewhere and is not under consideration at another journal. We have no conflicts of interest to declare.

We suggest the following potential reviewers with relevant expertise:

\begin{enumerate}
    \item \textbf{Benedetta Bolognesi} (Institute for Bioengineering of Catalonia) --- Expert in amyloid DMS, senior author of both DMS datasets used in this study
    \item \textbf{Bowen Jing} (MIT) --- Developer of AlphaFlow, expert in flow matching for protein ensembles
    \item \textbf{Alexander Rives} (EvolutionaryScale) --- Senior author of ESM-2, expert in protein language model scaling
\end{enumerate}

Thank you for considering our submission. We look forward to your response.

\closing{Sincerely,}

\end{letter}
\end{document}
